\LARGE\singlespacing{\textbf{S1 Appendix} \\ \medskip Transmission of respiratory viral infections with and without air cleaners in a Swiss school under non-pandemic conditions in 2023: A modeling study of epidemiological, environmental, and molecular data}

### 2023-08-22 doc/paper/manuscript.tex
Updates from Overleaf

\noindent\textbf{Methods and Findings:} We collected epidemiological (absences related to respiratory infections), environmental (CO$_2$ and particle concentrations), and molecular data (bioaerosol and saliva samples) over seven weeks from 16 January to 11 March 2023 in a secondary school (n=38 students in two classes) in Switzerland. We performed laboratory and molecular analyses to detect a panel of 19 respiratory viruses. We installed air cleaners in a cross-over design and compared positive samples and particle concentrations between weeks with and without air filters. We modeled the number of new infections using a Bayesian negative binomial regression model, adjusting for the incubation period, the number of recent and cumulative infections, the ventilation rate, the number of students in class, and community transmission. \\
Molecular analysis detected multiple viruses in saliva samples throughout the study (50/448 positive: 15~influenza B, 15~rhinovirus, 14~adenovirus, 3~SARS-CoV-2, 2~metapneumovirus, and 1~parainfluenza virus), but rarely in bioaerosol samples (2/105 positive: 1~rhinovirus and 1~adenovirus) and on the filters of air cleaners (4/160~positive: 1~influenza~B, 1~HRV, 1~adenovirus, and 1~SARS-CoV-2). The detection rate of positive saliva samples was comparable between study conditions (adjusted risk ratio 0.93, 95\%-Credible interval [CrI] 0.49 to 1.61). Daily average CO$_2$ levels were 1,769$\pm$391\,ppm ($\pm$ standard deviation) with air cleaners and 1,636$\pm$341\,ppm without air cleaners, corresponding to similar air change rates (0.74$\pm$0.41\,h$^{-1}$ vs. 0.68$\pm$0.32\,h$^{-1}$). Daily average aerosol number concentrations were 27$\pm$34\,1/cm$^3$ with and 95$\pm$81\,1/cm$^3$ without air cleaners, corresponding to an estimated adjusted decrease of 77\% (95\%-CrI 63\% to 86\%). Absences related to respiratory infections were less frequent with air cleaners in place (13~absences with vs. 21~absences without air cleaners). The estimated risk of infection tended to be lower with air cleaners (adjusted risk ratio 0.72, 95\%-Credible Interval 0.44 to 1.15). \medskip

@@

\noindent We computed the daily number of detected coughs and compared it between study conditions. We estimated the reduction in the number of coughs using a Bayesian Negative Binomial regression model (Supplementary Text~\zref{sec:aud-regression-model}) with time in class as the model offset, adjusting for class and weekday effects, the number of students in class, the air change rate, and the cumulative number of respiratory cases. \medskip

@@

We analyzed 448~saliva samples over the study across classes. We detected 15~influenza~B~(Flu B), 15~rhinoviruses~(HRV), 14~adenovirus (AdV), 3~SARS-CoV-2~(CoV), 2~metapneumoviruses~(MPV), and 1~parainfluenza~virus~(PIV) (\Cref{fig:molecular-descriptives}a). There was an equal number of positive samples with and without air cleaners (\Cref{fig:molecular-descriptives}b). Taking into account the number of samples and the number of suceptible students, the positivity rate remained comparable between study conditions (adjusted risk ratio 0.93, 95\%-Credible interval [CrI] 0.49 to 1.61), and was not sensitivity to infection to testing delays (Figure~\zref{fig:mol-estimation-results-sensitivity} in \supp). The distribution of positive samples varied between classes (\Cref{fig:molecular-descriptives}c). For example, in the first three study weeks all except for one sample were positive for AdV in class~A, while the vast majority of positive Flu~B samples were found in class~B. The virus-specific variations between classes are suggestive of transmission occurring within classes.

@@

Overall, there were 35~respiratory cases (absences related to respiratory infections) across classes. Incidentally, there were more cases in class~B (21) than in class~A (14), which is the class were more students recently recovered from a SARS-CoV-2 infection or were vaccinated. Except for one study week, there was an equal or higher number of cases without air cleaners (\Cref{fig:redcap-results}a). Based on our Bayesian latent hierarchical regression model, we estimated a trend towards a lower risk of infection with air cleaners (adjusted risk ratio 0.72, 95\%-Credible Interval 0.44 to 1.15). As a result, the estimated number of infections if air cleaners had been installed throughout the study is expected to be 36 (95\%-CrI 21 to 64) compared to 62 (95\%-CrI 29 to 141) infections if they would not have been installed (\Cref{fig:redcap-results}b). Detailed estimation results are provided in Supplementary Text~\zref{sec:detailed-redcap} in \supp.

\begin{figure}[!htpb]
    \includegraphics[width=\linewidth]{../../results/epi-data/cases_by_week.pdf} 
    \includegraphics[width=\linewidth]{../../results/epi-data/avoided-infections.pdf}
    \caption{\textbf{Analysis of epidemiological data and estimated relative risk of infection with air cleaners.}. \textbf{(a)}~Weekly number of respiratory cases with and without air cleaners. \textbf{(b)}~Estimated mean number of infections (posterior mean as dot and 50\%, 80\%, and 95\%-CrI as shaded areas) if, in both classes, air cleaners had been installed throughout the study or not been installed.}
    \label{fig:redcap-results}
\end{figure}

@@

Slightly fewer coughs were detected during study weeks where air cleaners were installed (\Cref{fig:coughing}). On average, there were 2.5~coughs/min with and 3.1~coughs/min without air cleaners (adjusted risk ratio 0.93, 95\%-CrI 0.85 to 1.02). Significantly more coughs were detected in class~B (adjusted risk ratio 2.16, 95\%-CrI 1.90 to 2.46), suggesting a link with the observed variation in virus-specific transmission between classes (\Cref{fig:molecular-descriptives}c).

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/cough-data/coughs-frequency-per-week.pdf}
    \caption{Daily average number of detected coughs per minute by study week, condition and class.}
    \label{fig:coughing}
\end{figure}

@@

\begin{figure}[!htpb]
    \includegraphics[width=\linewidth]{../../results/comparison.pdf} 
    \caption{\textbf{Comparison of the results between our previous study in 2022 during the COVID-19 pandemic (blue) and this study in 2023 during non-pandemic times (black).}. \textbf{(a)}~Proportion of positive saliva samples and virus. Viruses: Influenza B (Flu B), rhinovirus (HRV), adenovirus (AdV), SARS-CoV-2 (CoV), metapneumovirus (MPV), parainfluenza virus (PIV). \textbf{(b)}~Number of positive saliva, bioaerosol, and HEPA-filter samples. \textbf{(c)}~Estimated reduction in aerosol number and particle mass concentrations (CN in 1/cm$^3$) and particle mass concentration (PM for particles of sizes <1 to <10~$\mu$m, respectively in $\mu$gm$^{-3}$) with air cleaners (posterior mean as squares and 95\%-CrI as lines). \textbf{(d)}~Adjusted infection risk ratio (posterior mean as squares and 95\%-CrI as lines).}
    \label{fig:comparison}
\end{figure}

@@

Molecular transmission networks based on targeted sequencing or whole genome sequencing documented transmission in the classrooms. xxxx (influenza b, adeno, etc.). The low rate of positive bioaerosol samples also means that it is unclear to what extent students airborne transmission occurred in classrooms. It may be possible that students had rather little exposure to respiratory viruses in schools and acquired their infections elsewhere. However, the distribution of saliva samples between classes shows remarkable differences. Adenovirus was spreading entirely in class~A in the first three study weeks and only two infections with influenza~B occurred over the study. By contrast, influenza~B was spreading throughout the study in class~B and infections with adenovirus were only detected in the last study week. Adenovirus infections often tend to be mild\cite{Kunz2010CIDR} and less associated with coughing than influenza\cite{Ma2018RMV}, which is in line with the lower frequency of detected coughs in class~B. Altogether, the class-specific patterns in the spread of respiratory viruses indicate that transmission of respiratory infections may have occurred at least to some extent within the classrooms.

>>>>

\noindent\textbf{Methods and Findings:} We collected epidemiological (absences related to respiratory infections), environmental (CO$_2$ and particle concentrations), and molecular data (bioaerosol and saliva samples) over seven weeks from 16 January to 11 March 2023 in a secondary school (n=38 students in two classes) in Switzerland. We performed laboratory and molecular analyses to detect a panel of 19 respiratory viruses. We installed air cleaners in a cross-over design and compared positive samples and particle concentrations between weeks with and without air filters. We modeled the number of new infections using a Bayesian negative binomial regression model, adjusting for the incubation period, the number of recent and cumulative infections, the ventilation rate, the number of students in class, and community transmission. \\
Molecular analysis detected multiple viruses in saliva samples throughout the study (50/448 positive: 15~influenza B, 15~rhinovirus, 14~adenovirus, 3~SARS-CoV-2, 2~metapneumovirus, and 1~parainfluenza virus), but rarely in bioaerosol samples (2/105 positive: 1~rhinovirus and 1~adenovirus) and on the filters of air cleaners (4/160~positive: 1~influenza~B, 1~HRV, 1~adenovirus, and 1~SARS-CoV-2). The detection rate of positive saliva samples was comparable between study conditions (adjusted relative risk 0.93, 95\%-Credible interval [CrI] 0.49 to 1.61). Daily average CO$_2$ levels were 1,769$\pm$391\,ppm ($\pm$ standard deviation) with air cleaners and 1,636$\pm$341\,ppm without air cleaners, corresponding to similar air change rates (0.74$\pm$0.41\,h$^{-1}$ vs. 0.68$\pm$0.32\,h$^{-1}$). Daily average aerosol number concentrations were 27$\pm$34\,1/cm$^3$ with and 95$\pm$81\,1/cm$^3$ without air cleaners, corresponding to an estimated adjusted decrease of 77\% (95\%-CrI 63\% to 86\%). Absences related to respiratory infections were less frequent with air cleaners in place (13~absences with vs. 21~absences without air cleaners). The estimated risk of infection tended to be lower with air cleaners (adjusted relative risk 0.72, 95\%-Credible Interval 0.44 to 1.15). \medskip

@@

\noindent We computed the daily number of detected coughs and compared it between study conditions. We estimated the reduction in the number of coughs using a Bayesian Negative Binomial regression model (Supplementary Text~\zref{sec:aud-regression-model}) with time in class as the model offset, adjusting for class and weekday effects, the number of students in class, the air change rate, and the cumulative number of respiratory cases. In addition, we used a Bayesian Negative Binomial hierarchical (random effects) regression model to estimate the association between the number of coughs and the number of positive saliva test results for respiratory viruses. \medskip

@@

We analyzed 448~saliva samples over the study across classes. We detected 15~influenza~B~(Flu B), 15~rhinoviruses~(HRV), 14~adenovirus (AdV), 3~SARS-CoV-2~(CoV), 2~metapneumoviruses~(MPV), and 1~parainfluenza~virus~(PIV) (\Cref{fig:molecular-descriptives}a). There was an equal number of positive samples with and without air cleaners (\Cref{fig:molecular-descriptives}b). Taking into account the number of samples and the number of suceptible students, the positivity rate remained comparable between study conditions (adjusted relative risk 0.93, 95\%-Credible interval [CrI] 0.49 to 1.61), and was not sensitivity to infection to testing delays (Figure~\zref{fig:mol-estimation-results-sensitivity} in \supp). The distribution of positive samples varied between classes (\Cref{fig:molecular-descriptives}c). For example, in the first three study weeks all except for one sample were positive for AdV in class~A, while the vast majority of positive Flu~B samples were found in class~B. The virus-specific variations between classes are suggestive of transmission occurring within classes.

@@

Overall, there were 35~respiratory cases (absences related to respiratory infections) across classes. Incidentally, there were more cases in class~B (21) than in class~A (14), which is the class were more students recently recovered from a SARS-CoV-2 infection or were vaccinated. Except for one study week, there was an equal or higher number of cases without air cleaners (\Cref{fig:redcap-results}a). Based on our Bayesian latent hierarchical regression model, we estimated a trend towards a lower risk of infection with air cleaners (adjusted relative risk 0.72, 95\%-Credible Interval 0.44 to 1.15). As a result, the estimated number of infections if air cleaners had been installed throughout the study is expected to be 36 (95\%-CrI 21 to 64) compared to 62 (95\%-CrI 29 to 141) infections if they would not have been installed (\Cref{fig:redcap-results}b). Detailed estimation results are provided in Supplementary Text~\zref{sec:detailed-redcap} in \supp.

\begin{figure}[!htpb]
    \includegraphics[width=\linewidth]{../../results/epi-data/cases_by_week.pdf} 
    \includegraphics[width=\linewidth]{../../results/epi-data/avoided-infections.pdf}
    \caption{\textbf{Analysis of epidemiological data and estimated number of infections with and without air cleaners.}. \textbf{(a)}~Weekly number of respiratory cases with and without air cleaners. \textbf{(b)}~Estimated mean number of infections (posterior mean as dot and 50\%, 80\%, and 95\%-CrI as shaded areas) if, in both classes, air cleaners had been installed throughout the study or not been installed.}
    \label{fig:redcap-results}
\end{figure}

@@

Slightly fewer coughs were detected during study weeks where air cleaners were installed (\Cref{fig:coughing}a). On average, there were 2.5~coughs/min with and 3.1~coughs/min without air cleaners (adjusted relative risk 0.93, 95\%-CrI 0.85 to 1.02). Significantly more coughs were detected in class~B (adjusted relative risk 2.16, 95\%-CrI 1.90 to 2.46), suggesting a link with the observed variation in virus-specific transmission between classes (\Cref{fig:molecular-descriptives}c). This link can also be seen in the association between the frequency of coughing and the number of positive saliva tests for respiratory viruses (\Cref{fig:coughing}b). Coughing was more frequent for Flu~B (relative risk 1.25, 95\%-CrI 1.03 to 1.53) and slightly less frequent with AdV (0.87, 95\%-CrI 0.49 to 1.12).

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/cough-data/coughs-frequency-per-week.pdf}
    \includegraphics{results/cough-data/coughs-virus-association.pdf}
    \caption{\textbf{Analysis of coughing data.} \textbf{(a)}~Daily average number of detected coughs per minute by study week, condition and class. \textbf{(b)}~Estimated association (relative risk; posterior mean as dot, 50\%, 80\%, and 95\%-CrI as lines) between cough frequency and number of positive saliva test results by respiratory virus.}
    \label{fig:coughing}
\end{figure}

@@

\begin{figure}[!htpb]
    \includegraphics[width=\linewidth]{../../results/comparison.pdf} 
    \caption{\textbf{Comparison of the results between our previous study in 2022 during the COVID-19 pandemic (blue) and this study in 2023 during non-pandemic times (black).}. \textbf{(a)}~Proportion of positive saliva samples and virus. Viruses: Influenza B (Flu B), rhinovirus (HRV), adenovirus (AdV), SARS-CoV-2 (CoV), metapneumovirus (MPV), parainfluenza virus (PIV). \textbf{(b)}~Number of positive saliva, bioaerosol, and HEPA-filter samples. \textbf{(c)}~Estimated reduction in aerosol number and particle mass concentrations (CN in 1/cm$^3$) and particle mass concentration (PM for particles of sizes <1 to <10~$\mu$m, respectively in $\mu$gm$^{-3}$) with air cleaners (posterior mean as squares and 95\%-CrI as lines). \textbf{(d)}~Adjusted relative risk of infection (posterior mean as squares and 95\%-CrI as lines).}
    \label{fig:comparison}
\end{figure}

@@

Molecular transmission networks based on targeted sequencing or whole genome sequencing documented transmission in the classrooms. xxxx (influenza b, adeno, etc.). The low rate of positive bioaerosol samples also means that it is unclear to what extent students airborne transmission occurred in classrooms. It may be possible that students had rather little exposure to respiratory viruses in schools and acquired their infections elsewhere. However, the distribution of saliva samples between classes shows remarkable differences. Adenovirus was spreading entirely in class~A in the first three study weeks and only two infections with influenza~B occurred over the study. By contrast, influenza~B was spreading throughout the study in class~B and infections with adenovirus were only detected in the last study week. Adenovirus infections often tend to be mild\cite{Kunz2010CIDR} and less associated with coughing than influenza\cite{Ma2018RMV}, which is in line with the comparably lower frequency of detected coughs in class~B. Altogether, the class-specific patterns in the spread of respiratory viruses indicate that transmission of respiratory infections may have occurred at least to some extent within the classrooms.

### 2023-08-22 doc/paper/supplements.tex
Updates from Overleaf

@@

\begin{figure}[!htb]
\centering
    \includegraphics[width=10cm]{../../results/mol-data/model-results.pdf}
    \caption[Sensitivity analysis for the estimated effect of air cleaners on the positivity rate of saliva samples]{Sensitivity analysis for the estimated effect of air cleaners on the positivity rate of saliva samples. Shown is the adjusted relative risk ratio (posterior mean as dot and 50\%-, 80\%, and 95\%-CrIs as lines) for air cleaners by lag of the molecular test results.}
    \label{fig:mol-estimation-results-sensitivity}
\end{figure}

>>>>

In addition, we analyze the association between the daily number of coughs $Y$ and the number of positive saliva test resutls for respiratory viruses with a Negative Binomial hierarchical regression model
\begin{align*}
    Y | \beta_0, \theta_{[v]}, \theta, \tau &\sim \text{Negative-Binomial}(\mu,\phi) \\
    \log \mu &= \log T + \beta_0 + \sum_v \theta_{[v]} \cdot \text{Saliva}_v \\
    \beta_0 &\sim \text{Student-t}_5(0, 2.5) \\
    \theta_{[\text{Flu~B}]}, \dots, \theta_{[\text{PIV}]} &\sim \text{Normal}\left(\theta, \tau\right) \\
    \theta &\sim \text{Student-t}_5\left(0, \frac{2.5}{s_{v}}\right) \\
    \tau &\sim \text{Half-Normal}\left(0, 1\right) \\
    \frac{1}{\sqrt{\phi}} &\sim \text{Half-Normal}(0,1),
\end{align*}
where $\theta_{[v]}$ is the partially pooled estimate for respiratory virus $v \in$ (Flu~B, HRV, AdV, CoV, MPV, PIV), $\tau$ estimates variation between viruses, and $s_v$ is the average standard deviation across the count variables of the respiratory viruses.

@@

\begin{figure}[!htb]
\centering
    \includegraphics[width=10cm]{../../results/mol-data/model-results.pdf}
    \caption[Sensitivity analysis for the estimated effect of air cleaners on the positivity rate of saliva samples]{Sensitivity analysis for the estimated effect of air cleaners on the positivity rate of saliva samples. Shown is the adjusted relative risk (posterior mean as dot and 50\%-, 80\%, and 95\%-CrIs as lines) for air cleaners by lag of the molecular test results.}
    \label{fig:mol-estimation-results-sensitivity}
\end{figure}
